


## Setting up the workspace

```{r loading}
library(tidyverse)
library(dplyr)
library(ggplot2)
ds_salaries<- read_csv("C:/Nidheesh documents/MPS NU/Study Materials/Q3/ALY 6060 Decision Support and Business Intelligence/module 5/ds_salaries.csv")
```

### Inspecting my dataset

```{r}
str(ds_salaries)

colnames(ds_salaries)

unique(ds_salaries$job_title)

unique(ds_salaries$company_location)
```

I can already see two problems with my dataset: it includes jobs that are way outside of my wheelhouse and it includes data entries from all over the globe.

## Filtering my data

In order to filter this data down to information that is useful for my particular needs I'll need to remove all job titles that I am not qualified for and focus the data down to include only entries for companies that I would likely be applying to; namely, companies in the United States.

```{r filtering_analyst_sal}
analyst_sal <- filter(ds_salaries, job_title %in% c("Data Analyst", "Analytics Engineer","Data Engineer","ML Engineer","Applied Machine Learning Scientist"), company_location %in% c("US"))
```

```{r viewing_analyst_sal}
str(analyst_sal)

colnames(analyst_sal)

unique(analyst_sal$job_title)

unique(analyst_sal$company_location)
```

```{r visualizing_analyst_sal}
salavg <- analyst_sal %>%
  group_by(job_title) %>%
  summarise_at(vars(salary_in_usd), list(avg_salary = mean))

ggplot(salavg, aes(x=job_title, y=avg_salary, fill=job_title))+
  geom_bar(stat="Identity")+
  scale_y_continuous(labels = scales::comma)+
  labs(fill = "Designation", x = "", y = "Average Salary", title = "Average salary provided in the US categorized by designation")
```

## Filtering data down to entry level positions

Let's take a look at just entry level positions.

```{r filter_entry_avg_sal}
entry_avg_sal <- filter(analyst_sal, experience_level %in% c("EN"))

entrysalavg <- entry_avg_sal %>% 
  group_by(job_title) %>% 
  summarise_at(vars(salary_in_usd), list(avg_salary = mean))

ggplot(entrysalavg, aes(x=job_title, y=avg_salary, fill=job_title))+
  geom_bar(stat="Identity")+
  scale_y_continuous(labels = scales::comma)+
  labs(fill = "Designation", x = "", y = "Average Salary", title= "Entry level salaries provided in the US for the interested roles")
```



## A preview of salary progression

```{r}
mid_avg_sal <- filter(analyst_sal, experience_level %in% c("MI"))

midsalavg <- mid_avg_sal %>% 
  group_by(job_title) %>% 
  summarise_at(vars(salary_in_usd), list(avg_salary = mean))

ggplot(midsalavg, aes(x=job_title, y=avg_salary, fill=job_title))+
  geom_bar(stat="Identity")+
  scale_y_continuous(labels = scales::comma)+
  labs(fill = "Designation", x = "", y = "Average Salary", title="Mid level salaries provided in the US for the interested roles")
```

## Comparing entry level positions and mid level positions

```{r}
compare_avg_sal <- filter(analyst_sal, experience_level %in% c("EN", "MI"))

## Create a new table to compare entry and mid level position salaries ##

compsalavg <- compare_avg_sal %>% 
  group_by(job_title, experience_level) %>% 
  summarise_at(vars(salary_in_usd), list(avg_salary = mean))

## Visualize the data ##

ggplot(compsalavg, aes(x=job_title, y=avg_salary, fill=experience_level))+
  geom_bar(stat="Identity", position="dodge")+
  scale_y_continuous(labels = scales::comma)+
  geom_text(aes(label = scales::comma(avg_salary)), position = position_dodge(0.9), vjust=1)+
  labs(x = "", y = "Average Salary", labels= "Entry Level", "Mid Level", title="Salary differences categorized by Designation and Experience")+
  scale_fill_discrete(name = "Experience Level", labels = c("Entry Level", "Mid Level"))
```

There's an obvious outlier here.  Let's check our data entries and see what's going on.

```{r}
ggplot(compare_avg_sal, aes(x=job_title, fill=experience_level))+
  geom_bar(position="dodge")+
  geom_text(stat='count', aes(label=..count..), position = position_dodge(0.9), vjust=1)+
  labs(x = "", y = "Count", labels= "Entry Level", "Mid Level", title="Number of entries categorized by Designation and Experience in US")+
  scale_fill_discrete(name = "Experience Level", labels = c("Entry Level", "Mid Level"))
```

## Final analysis

```{r}
da_reliable <- filter(compsalavg, job_title %in% c("Data Engineer", "Data Analyst"))


ggplot(da_reliable, aes(x=job_title, y=avg_salary, fill=experience_level))+
  geom_bar(stat="Identity", position="dodge")+
  scale_y_continuous(labels = scales::comma)+
  geom_text(aes(label = scales::comma(avg_salary)), position = position_dodge(0.9), vjust=1)+
  labs(x = "", y = "Average Salary", labels= "Entry Level", "Mid Level", title="Comparison of subjectively acceptable salary data")+
  scale_fill_discrete(name = "Experience Level", labels = c("Entry Level", "Mid Level"))
```
